---
title: Decline in HRT Usage and Breast Cancer Incidence in 17 countries
author: Arindam Basu
output: 
  pdf_document: 
    keep_tex: yes
---

The purpose of this document is to describe the association between the decline in the  contains the R codes for the breast cancer study.

```{r}
## set the working directory

Mywd <- setwd("/Users/arindambose/GitHub/brca-study")

## Read the data into R

mydata <- read.csv("breastca.csv", header = TRUE, sep = ",")
mydata$hrtuse2 <- (mydata$declinehrtuse)^2
```

We regressed the decline in breast cancer incidence on the percentage decline in the usage of HRT in the countries studies. The 

```{r}

 plot(mydata$declinehrtuse, mydata$cadecrease,
     main = "Decrease in Breast Cancer Incidence with HRT",
    ylab = "Decrease in Breast Ca Incidence",
    xlab = "decrease in HRT usage")

abline(cahrt <- lm(mydata$cadecrease ~ mydata$declinehrtuse,  
                  data = mydata))

summary(cahrt)
```

As can be seen from the results, there is a small and statistically non-significant increment in the decline rates in breast cancer incidence (beta = 0.085, p = 0.45). In addition to this, we also get to see that for five countries, while HRT usage declined, this was not followed by corresponding decline in Breast cancer incidence in those countries. For these countries, there were increment in the incidence rates of breast cancer. 


```{r}
## You can use predict and lines() to create a quadratic plot
 ## first set up the quadratic terms
mydata$dechrtuse2 = mydata$declinehrtuse * mydata$declinehrtuse
# then we set up the model
#mydata1 <- mydata[,!is.na(mydata)]
quadmodel <- lm(cadecrease ~ declinehrtuse + dechrtuse2, data = mydata)
abline(quadmodel)

```


## You can use predict and lines() to create a quadratic plot
 ## first set up the quadratic terms
mydata$dechrtuse2 = mydata$declinehrtuse * mydata$declinehrtuse
# then we set up the model
# mydata1 <- mydata[,!is.na(mydata)]
#quadmodel <- lm(cadecrease ~ declinehrtuse + dechrtuse2, data = mydata)
#lines(mydata$declinehrtuse, predict(quadmodel))

###

## We remove data where the decrease in breast cancer were < 0

 mydata2 <- subset(mydata, cadecrease >= 0)

# print(summarize(mydata2))
 
# plot(mydata2$declinehrtuse, mydata2$cadecrease,
#          main = "Decrease in Breast Cancer Incidence with HRT",
#          ylab = "Decrease in Breast Ca Incidence",
#          xlab = "decrease in HRT usage")
 
# abline(lm(mydata2$cadecrease ~ mydata2$declinehrtuse, data = mydata2))

print(summary(mydata2$cadecrease))

# mydata2$highredca <- mydata2$cadecrease > 6.3

# print(aggregate(declinehrtuse ~ highredca, mydata2, mean))

mydata2$caredcat <- (cut(mydata2$cadecrease, breaks = c(0, 3.7, 10, 21)))
#mydata2$caredcat <- (mydata2$caredcat)
#levels(mydata2$caredcat) = c("leq 0", "0.1-3.2", "3.3-8.6", "8.7-21")
caredcatnames = c("small decrease", 
                "moderate decrease", "large decrease")
print(table(mydata2$caredcat))
hrtcadeclevel <- tapply(mydata2$declinehrtuse, mydata2$caredcat, mean)
print(hrtcadeclevel)

# barplot(hrtcadeclevel,
#        col = "black", main = "Bar Plot of % HRT Reduction",
#        xlab = "Extent of Decrease in Breast Cancer",
#        ylab = "Extent of Decrease in HRT Usage",
#        names.arg = caredcatnames,
#        ylim = c(0, 70))

## Now, analyse the the extent to reduction in HRT usage
##print(summary(mydata2$declinehrtuse))
## given the minimum 8, q1 21.8, q3 65.5, highest 75
mydata2$hrtdeclinecat <- cut(mydata2$declinehrtuse,
                            breaks = c(8, 30, 65.6, 75))
print(table(mydata2$hrtdeclinecat))
cadeclevel <- tapply(mydata2$cadecrease, mydata2$hrtdeclinecat, mean)
print(cadeclevel)

boxplot(mydata2$cadecrease ~ mydata2$hrtdeclinecat)

#barplot(cadeclevel,
 #       col = "black", main = "Bar Plot of Breast Ca Reduction",
  #      ylab = "Extent of Decrease in Breast Cancer",
  #      xlab = "Extent of Decrease in HRT Usage",
  #      names.arg = caredcatnames, srt = 45,
  #      ylim = c(0, 20))
